---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r reading data ,echo=TRUE}
data<-read.csv("~/RepData_PeerAssessment1/activity.csv")
```

## What is mean total number of steps taken per day?

```{r steps per day ,echo=TRUE}
library(dplyr)
by_date<- group_by(data,date)
steps_by_date<-summarise(by_date, dailysteps = sum(steps, na.rm = TRUE))
daily_total_steps<- as.vector(steps_by_date$dailysteps)
hist(daily_total_steps,main="Number of steps per day")
mean(daily_total_steps)
median(daily_total_steps)
```


## What is the average daily activity pattern?
```{r avtivity pattern ,echo=TRUE}
by_interval<- group_by(data,interval)
activity_pattern<-summarise(by_interval, interval_steps = mean(steps, na.rm = TRUE))
plot(activity_pattern,type="l",main="Average daily activity pattern")
highest<-arrange(activity_pattern,desc(interval_steps))
highest[1,]
```
## Imputing missing values
The stategy used for imputing missing values is to relpace NA's with the mean number of steps for each interval from the previous computation. 
```{r imputing missing values ,echo=TRUE}
nas<-is.na(data["steps"])
sum(nas)
merged_data<-merge(data,activity_pattern,all=TRUE)
merged_data_replaced<-mutate(merged_data,steps_relacement=ifelse(is.na(steps),interval_steps,steps))
new_data2<-select(merged_data_replaced,steps_relacement,date,interval)
new_data1<-rename(new_data2,steps=steps_relacement)
new_data<-arrange(new_data1,date,interval)
by_date<- group_by(new_data,date)
steps_by_date<-summarise(by_date, dailysteps = sum(steps, na.rm = TRUE))
daily_total_steps<- as.vector(steps_by_date$dailysteps)
hist(daily_total_steps,main="Number of steps per day")
mean(daily_total_steps)
median(daily_total_steps)
```
## Are there differences in activity patterns between weekdays and weekends?
```{r activity pattern weekday weekends,echo=TRUE}
library(lubridate)
new_data_with_day<-mutate(new_data,day=wday(date,label=FALSE),wday_wend=ifelse(day==1|day==7,"weekend","weekday"))
by_interval_day<- group_by(new_data_with_day,interval,wday_wend)
activity_pattern_day<-mutate(by_interval_day, steps_average = mean(steps, na.rm = TRUE))
library(lattice)
activity_pattern_day <- transform(activity_pattern_day, wday_wend = factor(wday_wend))
xyplot(steps_average ~ interval | wday_wend, data = activity_pattern_day, type="l",main="Activity patterns between weekdays and weekends",layout = c(1, 2))
```
